---
title: 'Assignment 3 - Final Report'
subtitle: 'Setting the environment'
author: 
  - Sharon Van Den Berg 9251936 (stvdb914@uowmail.edu.au) 
  - School of Mathematics and Applied Statistics
date: "`r format(Sys.time(), '%d %B, %Y \n')`"
knit: (function(input, encoding) {
  rmarkdown::render(input,
                    encoding = "UTF-8",
                    output_dir = "../report",
                    knit_root_dir = rprojroot::find_rstudio_root_file())})
output: 
  #html_document:
  bookdown::pdf_document2: 
     keep_tex: true
     number_sections: true
     toc: false
geometry: margin= 2.00cm
font-size: 11pt
font-style: "Serif"
line-spacing: 1.5
header-includes:
  - \usepackage{titling}
  - \pretitle{\begin{center}
    \includegraphics[width=6in,height=5in]{../images/UOW.png}\LARGE\\}
  - \posttitle{\end{center}}
  - \usepackage{setspace}
  - \onehalfspacing
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \lhead{\fontsize{11pt}{11pt}\selectfont "Times New Roman"}
  - \fancyhead[L]{\fontsize{11}{11}\selectfont STAT950/STAT250 ASSESSMENT 03 - Sharon Van Den Berg 9251936}
  - \fancyhead[R]{}
  - \lfoot{\fontsize{11pt}{11pt}\selectfont "Serif"}
  - \fancyfoot[C]{\thepage}
  - \fancyfoot[L]{}
  - \usepackage{pdflscape}
  - \newcommand{\blandscape}{\begin{landscape}}
  - \newcommand{\elandscape}{\end{landscape}}
documentclass: article
editor_options: 
  chunk_output_type: console
bibliography: DSAA811.bib
link-citations: TRUE
citation_package: apa-annotated.csl
---

**DECLARATION**
	No part of this Assignment has been copied from anyone else, and I have not lent any part of it to anyone else. No part of this assignment has been written by generative AI.

```{=tex}
	\vspace*{2em}\noindent
	\hfill%
	\begin{tabular}[t]{c}
		\includegraphics[width=3cm]{../images/Sharon.png} \\
		\rule{10em}{0.5pt}\\ Sharon Van Den Berg (9251936)
	\end{tabular}%
	\hfill%
	\begin{tabular}[t]{c}
		\today \\
		\rule{10em}{0.4pt}\\ Date
	\end{tabular}
```

[This report was made using R Markdown as a project linking to my Git account](https://github.com/stvdb914/DSAA811-Final-Report "Git Repository")

```{r child = './src/00_loader.Rmd' , eval = TRUE}
```

\newpage
# Abstract {-}
```{r child = './src/01_Abstract.Rmd'}
```

\newpage
\tableofcontents

# Glossary {-}
```{r Glossary, include = TRUE, echo = FALSE}
  options(knitr.kable.NA = '')

    glossaryDef %>% 
    knitr::kable("pipe", 
                 booktabs = TRUE) %>% 
    kable_styling(latex_options = "striped", position = "left") %>% 
      column_spec(1, bold = TRUE)
```

\newpage
# Introduction {-}

```{r child = './src/02_Introduction.Rmd'}
```

\newpage
# Background {-}
```{r child = './src/03_Background.Rmd'}
```

\newpage
# Research questions and aims of the project {-}
```{r child = './src/04_ResearchQuestions.Rmd'}
```

# Rationale {-}
```{r child = './src/05_Rational.Rmd'}
```

\newpage
# Data Description {-}
```{r child = './src/06_DataDescription.Rmd'}
```

\newpage
# Exploritory data analysis {-}
```{r child = './src/07_Exploritory.Rmd'}
```

\newpage
# Conclusion / Discussion {-}
```{r child = './src/08_Conclude.Rmd'}
```

\newpage
#Session Information {-}
```{r sessionInfo}
sessionInfo()
```

\newpage
# Bibliography {-}